---
title: "Quick-look diagnostics"
output: 
  html_document:
    toc: true
    toc_float: true
    depth: 2
params:
  run: 102
  modelDir: "model/pk"
  logDV: FALSE
  diagContCov: NULL
  diagCatCov: NULL
  excludeETAs: !r c() 
  script: "diagnostics-basic.Rmd"
  yspec: "data/spec/analysis3.yml"
  drugNameUnits: "concentration (mg/L)"
---

# Purpose

The main purpose of this template is to produce a set of quick diagnostic 
plots you can use during the early stages of model development to get a quick 
look at how your model fits the data. This template will only create an html 
preview of some typical diagnostic plots. It will read in covariates 
of interest from your spec file but it will not attempt to rename your ETAs or 
save out any plots. 

These plots are **not** intended to be report quality - please use the 
*diagnostics-report.Rmd* template to create diagnostics suitable for your report.

Please remember that these plots are just meant to provide a quick look at your
model diagnostics. They are not an exhaustive list of every possible plot 
and were chosen with this projects aims in mind. 

While this *should* give users examples of plots generated with the most up-to-date 
packages and methods, we're always happy to have feedback. If you know of more 
efficient methods or want to suggest alternative ways of plotting the figures 
please open an issue with the details. 


# Set up

```{r setup_general, include = FALSE, message=FALSE, warning = FALSE}

### General script set-up
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.height = 6, fig.width = 6)

### Libraries ----------------------------
library(tidyverse) 
library(pmplots)
library(bbr)
library(PKPDmisc)
library(yaml)
library(yspec)
library(mrggsave)
library(here) 
library(glue)
library(patchwork)
library(knitr)
library(pmtables)

### Options ----------------------------
options(mrg.script = params$script)
margin <- theme(plot.margin = unit(c(0.2, 0.2, 0.1, 0.1), "in"))
parens <-  pmtables::ensure_parens
```


```{r load_any_functions, include = FALSE}
### Source function scripts
source(here('script/functions-diagnostics.R'))
```


```{r setup_bbr, echo = FALSE, message = FALSE}
### Model location 
# Define `modelName` and path to this model directory

# Set directories for bbr models ----------------------------
modelName <- params$run
thisModelPath <- file.path(params$modelDir, modelName)
```


# Model details - Run number: `r params$run`

```{r modelDetails, echo = FALSE, results = "asis"}
mod <- read_model(here::here(thisModelPath))
mod
```

```{r modelOutputSummary, echo = FALSE, results = "asis"}
sum <- mod %>% model_summary()
sum
```

# Load Spec and define covariates
The covariates can be defined in several ways, please use whichever you find 
most convenient: 

1. You can pass them in through _params_ when you render the Rmd file.
2. You can define them using flags in your yspec (expected to be called 
    diagContCov and diagCatCov but you can update it).
3. You can define them directly in the Rmd code chunk below. 

Please note that the order described above will be preserved so, if for example, 
you choose to define the covariates with flags in your yspec, you can over ride
these choices by passing in a specific set of covariates when rendering the Rmd. 
Both option 1 and 2 would override the choice of covariates listed in the code 
below.


```{r yspec_details, echo = FALSE}
# load in the spec file
spec <- ys_load(here(params$yspec))

noContCov = FALSE
noCatCov = FALSE

# check for continuous covariate flags
if(!is_null(params$diagContCov)){
  # get covs from params list
  diagContCov = params$diagContCov
} else {
  # get covs from yspec
  diagContCov <- pull_meta(spec, "flags")[["diagContCov"]]
  if(is.null(diagContCov)) {
    # if no covs provided
    contCovMessage = "No continuous covariates were provided. To generate plots with continuous covariates please either (1) make a 'diagContCov' flag in your yspec file, (2) include covariates in the params section at the top of the 'diagnostics-basic.Rmd' file, or (3) pass covariates in through params$diagContCov when you render the Rmd"
    contCovMessageShort = "No continuous covariates were provided."
    noContCov = TRUE
  }
}

if(!is_null(params$diagCatCov)){
  # get covs from params list
  diagCatCov = params$diagCatCov
} else {
  # get covs from yspec
  diagCatCov <- pull_meta(spec, "flags")[["diagCatCov"]]
  if(is.null(diagCatCov)) {
    # if no covs provided
    catCovMessage = "No categorical covariates were provided. To generate plots with categoricall covariates please either (1) make a 'diagCatCov' flag in your yspec file, (2) include covariates in the params section at the top of the 'diagnostics-basic.Rmd' file, or (3) pass covariates in through params$diagCatCov when you render the Rmd"
    catCovMessageShort = "No categorical covariates were provided."
    noCatCov = TRUE
  }
}

```


# Read in data
Read in the model details using `read_model`. Details stored in the `mod` object 
can be used to identify the location of the source data (used in $DATA) 
- to see how this is done look at the `bbr::get_data_path()` and 
`bbr::build_path_from_model()` helper functions. 

After reading in the nonmem dataset and the output dataset they're joined by a 
`NUM` column. **This assumes that a row number column (called `NUM`) was 
included during data assembly.** The idea here is that in NONMEM, you table 
just `NUM` and none of the other input data items. They all will get joined 
back to the nonmem output ... even character columns.

The `data` used in the diagnostic plots has been filtered to only include the 
observations (i.e. `EVID==0`). Note that further modifications maybe needed, 
for example, if BLQ data was included in the model or if the `DV` was 
log-transformed. The dataset also converts the categorical covariates of 
interest to factors using the `yspec_add_factors` function and details 
described in the spec file.

The `id` subset gets the first record per ID. This would usually be the 
baseline value but consider filtering on a baseline flag if available. Also, if 
the model includes inter-occassion variaibility (IOV), the occassion variable 
should be included within the `distinct` function. 

```{r read_in_data, echo = FALSE, message=FALSE}
# Read in model
mod <- read_model(here::here(thisModelPath))

# Get original dataset and nonmem output
data0 <- nm_join(mod)

# filter to observation rows only
data <- data0 %>% 
  filter(EVID==0) 

# Add factors to the categorical covariates so the category names (not numbers)
# as shown on the plots. Note, if using a yspec version < 0.5.1 this 'all_of' option is 
# not available (you would need to manually type them in). You can either update your 
# yspec version or make plots with the numerical categories shown. 
if(packageVersion("yspec") >= "0.5.2"){
  data = data %>% 
    yspec_add_factors(spec, all_of(diagCatCov), .suffix = "")
}

# If log transformed DV used
if(isTRUE(params$logDV)) {
  if("DV.DATA" %in% names(data0)){
  data <- mutate(data, 
                 DV = exp(DV),
                 IPRED = exp(IPRED), PRED = exp(PRED))  
  } else {
    stop("To use logDV=TRUE you need to table out DV, PRED, IPRED in your NONMEM model.")
  }
}

# check for npdes - if none then plots will not be created
if("NPDE" %in% names(data)) {
  plotNPDE = TRUE
} else { plotNPDE = FALSE}

# identify the etas included in the dataset 
etas = str_subset(names(data), "ETA")

# the user can choose to exclude specific etas from eta plots, e.g., IOV etas
if(!is_null(params$excludeETAs)){
  etas = etas[which(!etas %in%  params$excludeETAs)]
} 

# filter to single row per ID and include factors 
id <- distinct(data, ID, .keep_all=TRUE) 
```


# General diagnostic plots

The following plots assume that the preferred x-axis labels are defined here. 

```{r xAxis_label_names, echo = FALSE}
### Ideally get units from yaml
xTIME <- glue(pm_axis_time(), xunit = parens(spec$TIME$unit))
xTAD <- glue(pm_axis_tad(), xunit = parens(spec$TAD$unit))
xPRED <-  glue(pm_axis_pred(), xname = params$drugNameUnits)
```

## DV vs PRED and IPRED 

Create plots of DV vs PRED and IPRED for the full dataset 

```{r DV_PRED_plots, echo = FALSE, message = FALSE}

dvp <- dv_pred(data, yname = params$drugNameUnits) + margin      # DV vs PRED 
dvip <- dv_ipred(data, yname = params$drugNameUnits) + margin    # DV vs IPRED 

p <- dvp / dvip
p
rm(p, dvp, dvip)
```



## NPDE plots

NPDE vs PRED, time and time after dose.

```{r NPDE_plots, eval=plotNPDE, echo = FALSE, message=FALSE}
##' Plan to display these three plots in a single, 3 panel figure and so the y-axis
##' labels are removed manually for display purposes
p1 <- npde_pred(data, x = xPRED, y = "NPDE //  ")
p2 <- npde_time(data, x = xTIME)
p3 <- npde_tad(data, x = xTAD, y = "NPDE //  ")
p <- p1 / p2 / p3 
p
rm(p, p1, p2, p3)
```


NPDE vs continuous covariates

```{r NPDE_cont_cov_plots, eval=plotNPDE, echo = FALSE, message = FALSE}
if(isFALSE(noContCov)){
  NPDEco <- spec %>% 
    ys_select(all_of(diagContCov)) %>%     # select the covariates of interest
    axis_col_labs(title_case = TRUE,   # converts yspec short to title_case 
                  short_max = 10) %>%  # if yspec short is >10 characters it keeps the abbreviation
    as.list()
  
  pList = purrr::map(NPDEco, ~ npde_cont(data, x = .x))
} else {
  pList = contCovMessage
}
pList
rm(pList)

```

NPDE vs categorical covariates.

```{r NPDE_cat_cov_plots, eval=plotNPDE, echo = FALSE, message = FALSE}
if(isFALSE(noCatCov)){
  NPDEca <- spec %>% 
    ys_select(all_of(diagCatCov)) %>%         # select the covariates of interest
    axis_col_labs(title_case = TRUE,   # converts yspec short to title_case 
                  short_max = 20) %>%  # if yspec short is >10 characters it keeps the abbreviation
    as.list()
  pList = purrr::map(NPDEca, ~ npde_cat(data, x = .x))
} else {
  pList = catCovMessage
}
pList
rm(pList)
```


## NPDE density histogram

```{r npde_hist_plots, eval=plotNPDE, echo = FALSE, message = FALSE}
p <- npde_hist(data)
p
rm(p)
```


## CWRES vs PRED, time and time after dose

```{r cwres_plots, echo = FALSE, message = FALSE}
p1 <- cwres_pred(data, x = xPRED, y = "CWRES //  ") 
p2 <- cwres_time(data, x = xTIME) 
p3 <- cwres_tad(data, x = xTAD, y = "CWRES //  ")
p = p1 / p2 / p3
p
rm(p, p1, p2, p3)
```



## CWRES qq and density plot

```{r qq_density_plots, echo = FALSE, message = FALSE}
# using patchwork to combine plots
p <- cwres_q(data) / cwres_hist(data)
p
rm(p)
```



# EBEs-based diagnostics

## ETA pairs plot 

```{r eta_pairs_plots, echo = FALSE, message = FALSE}
p <- eta_pairs(id, etas)
p
rm(p)
```


## Continuous covariate plots

```{r get_cont_cov_labels, echo = FALSE}
if(isFALSE(noContCov)){
  # These plots uses the yspec to automatically rename the axis labels.
  co <- spec %>% 
    ys_select(all_of(diagContCov)) %>%     # select the covariates of interest
    axis_col_labs(title_case = TRUE,  # converts yspec short to title_case 
                  short_max = 10)     # if yspec short is >10 characters it keeps the abbreviation
} 
```

### ETA vs continous covariates

Note that here we use a function that maps over the ETAs (not the covariates) because 
the purpose of these plots was to determine whether there were any trends in 
the covariates for a given ETA. This may need to be edited to address different study specific questions. 

```{r eta_vs_cont_plots, echo = FALSE, message = FALSE}
if(isFALSE(noContCov)){
  if (length(co) == 1) {
    p <- wrap_eta_cont(id, co, etas)
  } else {
    p <- purrr::map(.x = etas, ~ map_wrap_eta_cont(.x, co, id))
  }
} else {
  p = contCovMessageShort
}
p
rm(p)

```

### Continous covariate pairs plot

```{r contCov_pairs_plots, echo = FALSE, message = FALSE}
if(isFALSE(noContCov)){
  p <- eta_pairs(id, co)
} else {
   p = contCovMessageShort
}
  p
  rm(p)
```


## Categorical covariate plots

```{r get_cat_cov_labels, echo = FALSE}
if(isFALSE(noCatCov)){
  # These plots uses the yspec to automatically rename the axis labels.
  ca <- spec %>% 
    ys_select(all_of(diagCatCov)) %>%   # select the covariates of interest
    axis_col_labs(title_case=TRUE,  # converts yspec short to title_case 
                  short_max = 20)     # if yspec short is >10 characters it keeps the abbreviation
} 
```


### ETA vs categorical covariates

Note that here we use a function that maps over the covariates (not the ETAs) because 
the purpose of these plots was to determine whether there is any difference in the 
distribution of ETAs across studies, dosing groups and disease states.
This should be updated to reflect the questions you're trying to address.


```{r eta_vs_cat_plots, echo = FALSE, message = FALSE}
if(isFALSE(noCatCov)){
  if (length(etas) > 1) {
    p <- purrr::map(.x = ca, ~ map_eta_cat(.x, etas, id)) 
  } else {
    ## CHECK: depending on the labels, this may need to be changed 
    p <- purrr::map(.x = ca, ~ eta_cat(id, .x, etas) + rot_x(45)) 
  }
} else{
   p = catCovMessageShort
}
p
rm(p)
```


# Session details

It is considered good practice to include these details at the end of all rmd scripts

```{r details, include = TRUE}
Sys.getenv("AMI_NAME")
sessioninfo::session_info()
bbr::bbi_version()
```
