---
title: "Individual plots of DV, PRED and IPRED over time or TAD"
output: 
  html_document:
    toc: true
    toc_float: true
    depth: 2
params:
  run: 106
  modelDir: "model/pk"
  script: "id-dv-pred-plots.Rmd"
  yspec: "analysis3.yml"
  drugNameUnits: "Mockdralazine concentration (mg/L)"
  include_code: FALSE
  include_plots: FALSE
  run_mrggsave: TRUE
---

# Purpose

To produce a set of individual diagnostic plots that show the PRED, IPRED and DV 
versus TIME and TAD.

Please note these plots are just meant to provide an example of what could be created and how. 
They are not an exhaustive list of every possible plot and were chosen with the project 
aims in mind. 

While this _should_ give users examples of plots generated with the most up-to-date 
packages and methods, we're always happy to have feedback. If you know of more 
efficient methods or want to suggest alternative ways of plotting the figures 
please open an issue with the details. 


# Set up

```{r setup_general, include = F}
### General script set-up
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
  fig.height = 5, fig.width = 5)

### Libraries ----------------------------
suppressPackageStartupMessages(library(tidyverse))
library(pmplots)
library(bbr)
library(PKPDmisc)
library(yaml)
library(yspec)
library(mrggsave)
suppressPackageStartupMessages(library(glue))
suppressPackageStartupMessages(library(cowplot))

### Directories ----------------------------
figDir <- here::here("deliv", "figure")
dataDir <- here::here("data")
sourceDataDir <- file.path(dataDir, "source")
derivedDataDir <- file.path(dataDir, "derived")

if(!file.exists(figDir)) dir.create(figDir)

thisScript <- params$script


### Options ----------------------------
options(mrg.script = params$script)
options("mrggsave.dev" = "pdf")
margin <- theme(plot.margin = unit(c(0.2, 0.2, 0.1, 0.1), "in"))
parens <-  pmtables::ensure_parens
```


```{r load_any_functions, include = F}
### Source function scripts
# source(here::here('script/functions-diagnostics.R'))
```

### Model location
Define `modelName` and path to the model directory (`MODEL_DIR`).

```{r setup_bbr, echo = params$include_code, message = FALSE}
# Set directories for bbr models ----------------------------
# must go up an additional directory here as this script is located within the
# diagnostics-templates directory
MODEL_DIR <- params$modelDir
modelName <- params$run
thisModelPath <- file.path(MODEL_DIR, modelName)
```

### Figure location
If saving figures out to pdf, define where those pdfs should be saved to.
Here the figures are saved to ```deliv > figure > model_run_number```

```{r makeDir_mrggsave, echo = params$include_code, eval = params$run_mrggsave}
thisFigDir <- file.path(figDir, modelName)
if(!file.exists(thisFigDir)) dir.create(thisFigDir)

### set mrggsave figure directory here
options(mrggsave.dir = thisFigDir)
```


# Model details - Run number `r params$run`

```{r modelDetails, echo = params$include_code, results = "asis"}
mod <- read_model(here::here(thisModelPath))
mod
```

```{r modelOutputSummary, echo = params$include_code, results = "asis"}
sum <- mod %>% model_summary()
sum
```


# Load Spec
The aim is to use the information in the spec file to label the figures automatically.

```{r yspec_details, echo = params$include_code}

# Tell R where to find the yml
specLo <- here::here("data", "spec", params$yspec)

# load in the spec file
spec <- ys_load(specLo) %>% ys_namespace("plot")

```


# Read in data
Read in the model details using `read_model`. Details stored in the `mod` object can be 
used to identify the location of the source data (used in $DATA) - to see how this is done
look at the `bbr::get_data_path()` and `bbr::build_path_from_model()` helper functions. 

After reading in the nonmem dataset and the output dataset they're joined by a `NUM` column. **This assumes that a row number column (called `NUM`) was included during data assembly.** The idea here is that in NONMEM, you table just `NUM` and none of the other input data items. They all will get joined back to the nonmem output ... even character columns.

The `data` used in the diagnostic plots has been filtered to only include the observations 
(i.e. `EVID==0`). Note that further modifications maybe needed, for example, if BLQ data was 
included in the model or if the `DV` was log-transformed. The dataset also converts
the categorical covariates of interest to factors using the `yspec_add_factors` function 
and details described in the spec file.

The `id` subset gets the first record per ID. This would usually be the baseline value
but consider filtering on a baseline flag if available. Also, if the model includes
inter-occassion variaibility (IOV), the occassion variable should be included within the `distinct` function. 

```{r read_in_data, echo = params$include_code, message=FALSE}

# Read in model
mod <- read_model(here::here(thisModelPath))

# Get original dataset and nonmem output
data0 <- nm_join(mod)

# filter to observation rows only
data <- data0 %>% 
  filter(EVID==0) %>% 
  yspec_add_factors(spec, STUDY, CP, RF, DOSE, .suffix = "")

# filter to single row per ID and include factors 
# id <- distinct(data, ID, .keep_all=TRUE) 
```


# Individual plots

The following plots assume that the preferred x-axis labels are defined here. 

```{r xAxis_label_names, echo = params$include_code}
### Ideally get units from yaml 
xTIME <- glue(pm_axis_time(), xunit = parens(spec$TIME$unit))
xTAD <- glue(pm_axis_tad(), xunit = parens(spec$TAD$unit)) 
yLAB <- params$drugNameUnits
```

# DV, PRED and IPRED vs TIME

Create plots of DV, PRED and IPRED vs time (linear and log scale) for the full dataset 

```{r DV_PRED_IPRED_TIME_plots, echo = params$include_code, message = FALSE}

dv_pred_ipred_time_linear = data %>% 
  dv_pred_ipred(
    # facets = c("STUDY", "ID"), # option coming soon in pmplots
    id_per_plot = 9,
    ncol = 3,
    xlab = xTIME,
    ylab = yLAB,
    angle = 45,
    use_theme = theme_bw()
  ) 

dv_pred_ipred_time_log = data %>% 
  dv_pred_ipred(
    # facets = c("STUDY", "ID"), # option coming soon in pmplots
    id_per_plot = 9,
    ncol = 3,
    xlab = xTIME,
    ylab = yLAB,
    angle = 45,
    log_y = TRUE,
    use_theme = theme_bw()
  ) 

```

```{r include_DV_PRED_IPRED_TIME_plots, eval = params$include_plots, include = params$include_plots, echo = F, message = FALSE}
print("DV, PRED and IPRED vs time (linear)")
dv_pred_ipred_time_linear
print("DV, PRED and IPRED vs time (log)")
dv_pred_ipred_time_log

```

```{r save_DV_PRED_IPRED_TIME_plots, include = FALSE, eval = params$run_mrggsave, message = FALSE}
# the stem can be made using c() or glue syntax 
mrggsave(dv_pred_ipred_time_linear, 
         stem = "id-dv-pred-ipred-time-linear", 
         width = 6, height = 7)

mrggsave(dv_pred_ipred_time_log, 
         stem = "id-dv-pred-ipred-time-log",
         width = 6, height = 7)


```

```{r rm_DV_PRED_IPRED_TIME_plots, include = F}
rm(dv_pred_ipred_time_linear, dv_pred_ipred_time_log)
```


# DV, PRED and IPRED vs TAD

Create plots of DV, PRED and IPRED vs TAD (linear and log scale) for the full dataset 

```{r DV_PRED_IPRED_TAD_plots, echo = params$include_code, message = FALSE}

dv_pred_ipred_tad_linear = data %>% 
  dv_pred_ipred(
    # facets = c("STUDY", "ID"), # option coming soon in pmplots
    x = pm_axis_tad(),
    id_per_plot = 9,
    ncol = 3,
    xlab = xTAD,
    ylab = yLAB,
    angle = 45,
    use_theme = theme_bw()
  ) 

dv_pred_ipred_tad_log = data %>% 
  dv_pred_ipred(
    # facets = c("STUDY", "ID"), # option coming soon in pmplots
    x = pm_axis_tad(),
    id_per_plot = 9,
    ncol = 3,
    xlab = xTAD,
    ylab = yLAB,
    angle = 45,
    log_y = TRUE,
    use_theme = theme_bw()
  ) 

```

```{r include_DV_PRED_IPRED_TAD_plots, eval = params$include_plots, include = params$include_plots, echo = F, message = FALSE}
print("DV, PRED and IPRED vs tad (linear)")
dv_pred_ipred_tad_linear
print("DV, PRED and IPRED vs tad (log)")
dv_pred_ipred_tad_log

```

```{r save_DV_PRED_IPRED_TAD_plots, include = FALSE, eval = params$run_mrggsave, message = FALSE}
# the stem can be made using c() or glue syntax 
mrggsave(dv_pred_ipred_tad_linear, 
         stem = "id-dv-pred-ipred-tad-linear", 
         width = 6, height = 7)

mrggsave(dv_pred_ipred_tad_log, 
         stem = "id-dv-pred-ipred-tad-log",
         width = 6, height = 7)


```

```{r rm_DV_PRED_IPRED_TAD_plots, include = F}
rm(dv_pred_ipred_tad_linear, dv_pred_ipred_tad_log)
```




# Session details

It is considered good practice to include these details at the end of all rmd scripts

```{r details, include = TRUE}
Sys.getenv("AMI_NAME")
sessioninfo::session_info()
bbr::bbi_version()
```


